---
title: "Dataset Manipulation Ouiz"
output: html_document
date: "2025-10-30"
---
---
title: "Quiz 2"
author: "Amanda Mae Woodward"
date: "2025-10-15"
output: html_document
---

This quiz will focus on data frame manipulation, however, you may need to rely on skills you've learned in prior weeks. 

1. load the starwars dataset from tidyverse
```{r}
library(tidyverse)
data("starwars")
```

2. separate the films column so that there is one film in each column (don't worry about the "c(" and ) at the end for now)
```{r}
starwars<-separate(starwars, films, into = c("Film1", "Film2", "Film3", "Film4", "Film5", "Film6", "Film7"), sep = ",")
```

3. alter the data from wide data to long data
```{r}
starwars<-separate(starwars, vehicles, into = c("Vehicle1", "Vehicle2"), sep = ",")
starwars<-separate(starwars, starships, into = c("Starship1", "Starship2", "Starship3", "Starship4", "Starship5"), sep = ",")
starwarsLong<-pivot_longer(starwars, cols = c("hair_color", "skin_color", "eye_color", "sex", "gender", "homeworld", "species", "Film1", "Film2", "Film3", "Film4", "Film5", "Film6", "Film7", "Vehicle1", "Vehicle2", "Starship1", "Starship2", "Starship3", "Starship4", "Starship5"), names_to = "NameOfDemographic", values_to = "DemographicInformation")
```

4. in the column where you stored the films, rename the levels so each film name is consistent (e.g. get rid of the "c(" and ")" )
```{r}
starwarsLong$DemographicInformation <- gsub('["()]|c\\(', '', starwarsLong$DemographicInformation)
```
I found this code on DigitalOcean which is a forum where someone mentioned that this could remove that information. 


5. Create a column in your dataset that indicates whether a character was in the first starwars film, *Star Wars Episode IV: A New Hope* 
```{r}
starwarsLong$NewHopeAppearance<-NA
starwarsLong$NewHopeAppearance<-ifelse(starwarsLong$DemographicInformation=="A New Hope", "Yes", "No")
```

6. BMI is calculated as mass in Kg/height in meters squared. You currently have mass in Kg and height in cm. Create a column that contains each character's BMI. 
```{r}
starwarsLong$BMI<-NA
starwarsLong<-starwarsLong %>% mutate(BMI = mass/sqrt((height/100)))
```

7. create a dataframe that contains only the name, hair color, and eye color for each character. 
```{r}
SimplifiedDataSW<- data.frame(starwars$name, starwars$hair_color, starwars$eye_color)
print(SimplifiedDataSW)
```

