---
title: "Dataset Manipulation Ouiz"
output: html_document
date: "2025-10-30"
---
---
title: "Quiz 2"
author: "Amanda Mae Woodward"
date: "2025-10-15"
output: html_document
---

This quiz will focus on data frame manipulation, however, you may need to rely on skills you've learned in prior weeks. 

1. load the starwars dataset from tidyverse
```{r}
library(tidyverse)
data("starwars")
```

2. separate the films column so that there is one film in each column (don't worry about the "c(" and ) at the end for now)
```{r}
starwars<-separate(starwars, films, into = c("Film1", "Film2", "Film3", "Film4", "Film5", "Film6", "Film7"), sep = ",")
```

3. alter the data from wide data to long data
```{r}
starwarsLong<-pivot_longer(starwars, cols = c(2:11), names_to = c("DemographicInfo"), values_to = "DemographicInfo")
```

4. in the column where you stored the films, rename the levels so each film name is consistent (e.g. get rid of the "c(" and ")" )
```{r}

```

5. Create a column in your dataset that indicates whether a character was in the first starwars film, *Star Wars Episode IV: A New Hope* 
```{r}

```

6. BMI is calculated as mass in Kg/height in meters squared. You currently have mass in Kg and height in cm. Create a column that contains each character's BMI. 
```{r}

```

7. create a dataframe that contains only the name, hair color, and eye color for each character. 
```{r}

```

