---
title: "Data Manipulation Assignment"
output: html_document
date: "2025-10-16"
---
---
title: "Dataset Manipulation Assignment"
author: "Amanda Mae Woodward"
date: "2025-10-07"
output: html_document
---

#Part 1
1.Open the sleep dataset in R 
```{r}
data(sleep)
```

2. Is it long or wide format? How do you know? 

It is in long format because each subject has at least two lines that pertain to data related to them, specifically the increase in hours of sleep depending upon which drug they took. 

3. Transform this data so that it is in the other format. 
```{r}
library(tidyverse)
sleepWide <- pivot_wider(sleep, id_cols = c("ID"), names_from = "group", values_from = "extra")
```

4. Transform your data back to its original format.
```{r}
sleepLong <- pivot_longer(sleepWide, cols = c("1", "2"), names_to = "Drug Given", values_to = "Sleep Change")
```

5. Open the Starwars dataset
```{r}
data(starwars)
```

6. Create homeworld as a factor. How many levels does it have? 
```{r}
class(starwars$homeworld)
starwars$homeworldFactor<- as.factor(starwars$homeworld)
starwars$homeworldFactor
```
It has 48 levels. 

7. Reorder homeworld so that Naboo is the reference group. 
```{r}
starwars$homeworldFactor<- relevel(starwars$homeworldFactor, ref = "Naboo")
starwars$homeworldFactor
```

8. Use mutate to create a column transforming kg to pounds. (hint: you may need to look up the conversion)
```{r}
starwars %>% mutate(pounds = mass/2.20462)
```

9. Print a summary of your column from #8
```{r}
summary(starwars$mass)
```

#Part 2
10. open the billboard dataset.
```{r}
data(billboard)
```

11. are the data in this dataframe in long or wide format? explain how you know. 

The data in this dataframe is in wide format. I can tell this because the information pertaining to the subject is on one row rather than in multiple rows. 

12. change the data from its original format to the other version. 
```{r}
billboardLong <- pivot_longer(billboard, cols = c(4:79), names_to = "Week", values_to = "Ranking")
```

13. change it from your altered version in #3 back to its original state. 

My assumption is it is meant to be the original state from #12? So that is what this is:
```{r}
billboardWide<-pivot_wider(billboardLong, id_cols = c("artist", "track", "date.entered"), names_from = "Week", values_from = "Ranking")
```

14. Separate the date entered column into three columns: one for year, one for month, and one for day. 
```{r}
separate(billboardWide, date.entered, into = c("Year", "Month", "Day"), sep = "-")
```
I found the solution of using the "-" to separate each part of the date by looking at a forum on posit where someone asked about separating dates, and one of the commenters mentioned trying that, so I did and it worked!
