---
title: "Foundations in R pt 2 Assignment"
output: html_document
date: "2025-10-02"
---
---
title: "foundations in R pt 2 assignment"
author: "Amanda Mae Woodward"
date: "2025-09-25"
output: html_document
---
1. Load the tidyverse library. 
```{r}
library(tidyverse)
```

2. Load the billboard data. 
```{r}
data("billboard")
```

3. What type of data is this? (e.g. describe what's in the dataset).

The data set is the song rankings for the Billboard Top 100 in the year 2000. Included in the data set are the artist name, the song name, the date that the song entered the top 100, and the rank of the song each week after it entered. I found this information by using the question mark function and writing billboard after it in the console. 

4. Print the column names from this dataset. 
```{r}
colnames(billboard)
```

5. Print the first 5 entries in the dataset. 
```{r}
head(billboard, 5)
```

# Practice
1. Create a subset of data that has Jay-Z as the only artist. How many songs are in it?
```{r}
JustJayZ<-subset(billboard, billboard$artist == "Jay-Z")
print(JustJayZ)
```
There are five songs in the subset. 

2. In the original dataset, use filter to find information about the song "All the small things." What was the highest position (i.e. lowest number) this song reached on the billboard 100? 
```{r}
SmallThings<- billboard %>% filter(track=="All The Small Things")
SmallThings
```
The highest position it achieved on Billboard's Top 100 was 6. 

3. Create a subset of data for songs Tim McGraw sang. Summarize the data for wk 1.
```{r}
JustMcGraw<-subset(billboard, billboard$artist == "McGraw, Tim")
summary(JustMcGraw$wk1)
```

4. Create a function that can be used to determine if Britney Spears sang the song. 
```{r}
BritneySong<-function(thing){thing == "Spears, Britney"}
```

5. Test your function on the billboard data. Save the results to a new column called "wasItBritney"
```{r}
billboard$wasitBritney<- NA
billboard$wasitBritney<-BritneySong(billboard$artist)
```

6. Use a for loop to determine if a song placed 50 or below or placed above 50 during the first week. 
```{r}
billboard$top50<- NA
for (i in 1:nrow(billboard)) {if(billboard$wk1[i]<50){billboard$top50[i]<-"Top 50"}else{billboard$top50[i]<-"Bottom 50"}}
```

7. Use an if else statement to determine the same information as you obtained with the for loop. 
```{r}
billboard$top50Part2<-ifelse(billboard$wk1<50, "Top 50", "Bottom 50")
```

8. Compare your results. 
```{r}
print(billboard$top50) 
print(billboard$top50Part2)
```

