---
title: "Foundations in R pt 2 Assignment"
output: html_document
date: "2025-10-02"
---
---
title: "foundations in R pt 2 assignment"
author: "Amanda Mae Woodward"
date: "2025-09-25"
output: html_document
---
1. Load the tidyverse library. 
```{r}
library(tidyverse)
```

2. Load the billboard data. 
```{r}
data("billboard")
```

3. What type of data is this? (e.g. describe what's in the dataset).



4. Print the column names from this dataset. 
```{r}

```

5. Print the first 5 entries in the dataset. 
```{r}

```


# Practice
1. Create a subset of data that has Jay-Z as the only artist. How many songs are in it?
```{r}

```

2. In the original dataset, use filter to find information about the song "All the small things." What was the highest position (i.e. lowest number) this song reached on the billboard 100? 
```{r}

```

3. Create a subset of data for songs Tim McGraw sang. Summarize the data for wk 1.
```{r}

```

4. Create a function that can be used to determine if Britney Spears sang the song. 
```{r}

```

5. Test your function on the billboard data. Save the results to a new column called "wasItBritney"
```{r}

```

6. Use a for loop to determine if a song placed 50 or below or placed above 50 during the first week. 
```{r}

```

7. Use an if else statement to determine the same information as you obtained with the for loop. 
```{r}

```

8. Compare your results. 
```{r}

```

