---
title: "GG Plot Assignment"
output: html_document
date: "2025-10-23"
---
---
title: "Ggplot part 1 assignment"
author: "Amanda Mae Woodward"
date: "2025-10-23"
output: html_document
---

1. open the starwars dataset in tidyverse
```{r}
library(tidyverse)
data(starwars)
```

2. create a graph showing the frequency of eye colors in the dataset. 
```{r}
ggplot(starwars, aes(eye_color))+geom_bar()
```

3. Customize your graph to make it more presentable. Consider color, backgrounds, axes, etc. 
```{r}
ggplot(starwars, aes(eye_color))+geom_bar(color = "steelblue", fill = "pink3")+theme_bw()+xlab("Eye Color")+ylab("Frequency")+ggtitle("Frequency of Eye Color in Star Wars Dataset")+theme(axis.text.x = element_text(angle = 90)) 
```
I found how to rotate the axis label to make them more readable on geeksforgeeks.org. 

4. Create a graph displaying information about character's height in each homeworld. 
```{r}
ggplot(starwars, aes(homeworld, height))+geom_point()
```

5. Customize your graph to make it more presentable. Consider color, backgrounds, axes, etc.
```{r}
ggplot(starwars, aes(homeworld, height, color = height))+geom_point()+theme_bw()+xlab("Homeworld")+ylab("Height (cm)")+ggtitle("Character's Height in Each Homeworld")+theme(axis.text.x = element_text(angle = 90))
```

6. Create a graph of character's height by mass. 
```{r}
ggplot(starwars, aes(mass, height))+geom_point()
```

7.Customize your graph to make it more presentable. Consider color, backgrounds, axes, etc.
```{r}
ggplot(starwars, aes(mass, height, color = height))+geom_point()+theme_bw()+xlab("Mass (kg)")+ylab("Height (cm)")+ggtitle("Character's Height by Mass")+guides(color = guide_legend(title = "Height"))
```

8. Add gender as a third variable to your graph of height by mass. 
```{r}
ggplot(starwars, aes(mass, height, color = gender))+geom_point()
```

9. Customize your graph to make it more presentable. Consider color, backgrounds, axes, etc.
```{r}
ggplot(starwars, aes(mass, height, color = gender))+geom_point()+theme_bw()+xlab("Mass (kg)")+ylab("Height (cm)")+ggtitle("Character's Height by Mass")+guides(color = guide_legend(title = "Gender"))
```
