---
title: "GG Plot Assignment Part 2"
output: html_document
date: "2025-10-30"
---

---
title: "ggplot part 2 assignment"
author: "Amanda Mae Woodward"
date: "2025-10-24"
output: html_document
---

1.  Open the starwars dataset (located in tidyverse)

```{r}
library(tidyverse)
data("starwars")
```

2.  Create a scatter plot of character's height and mass.

```{r}
ggplot(starwars, aes(height, mass))+geom_point()+theme_bw()
```

2a. Change the color and shape of the points in your scatter plot.

```{r}
ggplot(starwars, aes(height, mass))+geom_point(color = "steelblue", shape = 4)+theme_bw()
```

3.  Add Homeworld as a third variable to your scatterplot.

```{r}
ggplot(starwars, aes(height, mass, color = homeworld))+geom_point(shape = 4)+theme_bw()
```

3a. What do you notice about the graph? (e.g. describe what you see)

The graph has a fairly consistent trend of the taller the character gets, the heavier their mass, except for one character that is a strong outlier (about 175 cm and over 1000 kg). It also appears that quite a few of the data points fall within a small range of a little above 150 cm to a little under 200 cm. 

4.  Make a boxplot of height.

```{r}
ggplot(starwars, aes(height))+geom_boxplot()
```

5.  Make a boxplot of height for each species present in the data.

```{r}
ggplot(starwars, aes(species, height))+geom_boxplot()
```

5a. change the appearance of your graph in at least 3 ways.

```{r}
ggplot(starwars, aes(species, height, color = species))+geom_boxplot()+theme_bw()+xlab("Species")+ylab("Height (cm)")+ggtitle("Character's Height by Species")+theme(axis.text.x = element_text(angle = 90))
```
I found how to tilt the labels on the x-axis on either the posit forum or the geeksforgeeks website, though I forget which. I also used this in the first ggplot assignment. 

6.  Using any variables in the starwars dataset, create a graph that uses a color palette from viridis.

```{r}
library(viridis)
ggplot(starwars, aes(mass, fill = homeworld))+ geom_histogram()+scale_fill_viridis(discrete = TRUE, option = "A")+theme_bw()
```

7.Using any variables in the starwars dataset, create a graph that uses a color palette from RColorBrewer.

```{r}
ggplot(starwars, aes(homeworld, species, color = sex))+geom_point()+scale_color_brewer(palette = "RdBu")+theme_bw()+theme(axis.text.x = element_text(angle = 90))
```

8.  Using any variables in the starwars dataset, create a graph using facet_wrap

```{r}
ggplot(starwars, aes(mass, height, color = species))+geom_point()+theme_bw()+facet_wrap(~sex)
```
