---
title: "PSY 4802 Quiz 1"
output: html_document
date: "2025-10-14"
---
---
title: "Quiz 1"
author: "Amanda Mae Woodward"
date: "2025-10-07"
output: html_document
---

This quiz will cover everything we've covered in foundations of R assignments.

1. Open the tidyverse library.
```{r}
library(tidyverse)
```

2. Open the starwars dataset.
```{r}
data("starwars")
```

3. Create a subset of data that contains only the humans from the starwars franchise.
```{r}
OnlyHuman<-subset(starwars, starwars$species == "Human")
OnlyHuman
```

4. In this subset of data, who is the tallest character? 
```{r}
summary(OnlyHuman$height)
print(OnlyHuman %>% filter(height=="202"))
```
The tallest height is 202 cm, which is Darth Vader.

5. In this subset of data, print the information about the character in the third row. 
```{r}
OnlyHuman[3, ]
```

6. In the full starwars dataset, calculate the average mass for the masculine characters. 
```{r}
MassMascSW <- starwars %>% filter(gender == "masculine") 
mean(MassMascSW$mass, na.rm = TRUE)
```
R does not appreciate looking at the average mass for the masculine characters, because there are some characters that do not have a mass listed. I searched online to find a way to exclude values that are NA, and I found the na.rm function through a few different websites, one being GeeksforGeeks.org, and I used that here. 

7. Create a dataset that contains only the characters and the films that they appeared in from the full dataset. 
```{r}
CharacterAndMovie<- starwars %>% select(name, films)
CharacterAndMovie
```

8. Create a new column in the starwars dataset that contains a "ranking" of all characters (Note: you don't have to actually rank them. Just fill the column in with the numbers 1-87)
```{r}
starwars$Ranking<- sample(1:87, 87, replace = TRUE)
print(starwars$Ranking)
```

9. Create a subset of data for the characters from Naboo. How many characters are from Naboo? 
```{r}
FromNaboo<-subset(starwars, starwars$homeworld == "Naboo")
FromNaboo
nrow(FromNaboo)
```
There are 11 characters from Naboo. 

10. How many droids are in the full dataset? 
```{r}
DroidOnly<-starwars[starwars$species=="Droid", ]
nrow(DroidOnly)
```
There are 10 droids in the full dataset. 
