---
title: "PSY 4802 Quiz 1"
output: html_document
date: "2025-10-14"
---
---
title: "Quiz 1"
author: "Amanda Mae Woodward"
date: "2025-10-07"
output: html_document
---

This quiz will cover everything we've covered in foundations of R assignments.

1. Open the tidyverse library.
```{r}
library(tidyverse)
```

2. Open the starwars dataset.
```{r}
data("starwars")
```

3. Create a subset of data that contains only the humans from the starwars franchise.
```{r}
OnlyHuman<-subset()
```

4. In this subset of data, who is the tallest character? 
```{r}

```

5. In this subset of data, print the information about the character in the third row. 
```{r}

```

6. In the full starwars dataset, calculate the average mass for the masculine characters. 
```{r}

```

7. Create a dataset that contains only the characters and the films that they appeared in from the full dataset. 
```{r}

```

8. Create a new column in the starwars dataset that contains a "ranking" of all characters (Note: you don't have to actually rank them. Just fill the column in with the numbers 1-87)
```{r}
CharacterRanking<-
```

9. Create a subset of data for the characters from Naboo. How many characters are from Naboo? 
```{r}

```

10. How many droids are in the full dataset? 
```{r}

```

